Trans-dimensional Bayesian Inference and Nested Sampling

### 2015-02-02 Documents/Talk/content.tex
More work

\begin{itemize}
\item Start at some point $\theta$ in the hypothesis space.
\item Loop\\
$\{$
  \begin{itemize}
  \item Generate {\bf proposal} from some distribution $q(\theta' | \theta)$
  (e.g. slightly perturb the current position).
  \item With probability$^*$ $\alpha = \min\left(1, \frac{p(\theta')p(D|\theta')}{p(\theta)p(D|\theta)}\right)$, accept the proposal (i.e. replace $\theta$ with $\theta'$).
  \item Otherwise, stay in the same place.
  \end{itemize}
$\}$
\end{itemize}
\end{frame}

@@

@@

{\bf Nested Sampling solves the problem of phase transitions}.
\end{frame}

\begin{frame}[t]{Diffusive Nested Sampling}
\begin{columns}[T]
\begin{column}{0.35\textwidth}
  \vspace{30pt}
  \begin{itemize}
  \setlength{\itemsep}{20pt}
  \item Replace the posterior distribution with a {\bf mixture of constrained priors}
  \item Can cross valleys
  \item Posterior is available via re-weighting
  \end{itemize}
\end{column}
\hfill
\begin{column}{0.5\textwidth}
  \hspace{50pt}
  \includegraphics[scale=0.4]{dnest.png}
\end{column}
\end{columns}
\end{frame}

\begin{frame}[t]{Diffusive Nested Sampling}
\begin{itemize}
\setlength{\itemsep}{20pt}
\item I've implemented the sinewave problem and the galaxy problem
\item The common features (what proposals to use?) are delegated to
a separate library called {\tt RJObject}
\item Uses birth-death MCMC, but exploring the DNS distribution, rather than
the (harder) posterior.
\end{itemize}
\end{frame}

@@

\begin{frame}[t]{Sinusoid data movie}
{\tt https://www.youtube.com/watch?v=q4s8NJyUVoY}
\end{frame}

@@

>>>>

\begin{itemize}
\item Start at some point $\theta$ in the hypothesis space.
\item Loop\\
$\{$
  \begin{itemize}
  \item Generate {\bf proposal} from some {\it proposal distribution} $q(\theta' | \theta)$
  (e.g. slightly perturb the current position).
  \item With probability $\alpha = \min\left(1, \frac{q(\theta | \theta')}{q(\theta' | \theta)}\frac{p(\theta')p(D|\theta')}{p(\theta)p(D|\theta)}\right)$, accept the proposal (i.e. replace $\theta$ with $\theta'$).
  \item Otherwise, stay in the same place.
  \end{itemize}
$\}$
\end{itemize}
\end{frame}

@@

\begin{frame}[t]{Alternative: Birth and Death}
Birth and death moves are surprisingly easy to implement. Design them to
explore the prior (\& satisfy detailed balance), and then use
\begin{eqnarray}
\alpha = \min\left(1, \frac{p(D|\theta')}{p(D|\theta)}\right)
\end{eqnarray}
as the acceptance probability: i.e. just the ratio of likelihoods.
\end{frame}

\begin{frame}[t]{Alternative: Birth and Death}
Birth proposal: Increment $N$, generate the new 'object' from the prior
(given current values of hyperparameters)

\vspace{20pt}

Death proposal: Choose one of the $N$ objects, remove it, decrement $N$.
\end{frame}

@@

{\bf Nested Sampling solves the problem of phase transitions and (as far as I
know) no other Monte Carlo method does.}
\end{frame}

\begin{frame}[t]{Diffusive Nested Sampling}
\begin{columns}[T]
\begin{column}{0.35\textwidth}
  \vspace{30pt}
  \begin{itemize}
  \setlength{\itemsep}{20pt}
  \item Replace the posterior distribution with a {\bf mixture of constrained priors}
  \item Can cross valleys (the prior is a mixture component)
  \item Posterior is available via re-weighting
  \end{itemize}
\end{column}
\hfill
\begin{column}{0.5\textwidth}
  \hspace{50pt}
  \includegraphics[scale=0.4]{dnest.png}
\end{column}
\end{columns}
\end{frame}

\begin{frame}[t]{Diffusive Nested Sampling}
\begin{itemize}
\setlength{\itemsep}{20pt}
\item I've implemented a library, {\tt RJObject}, that handles all the
Metropolis proposals I need for these kinds of problems.
\item Uses birth-death MCMC, but exploring the DNS distribution, rather than
the (harder) posterior.
\end{itemize}
\end{frame}

@@

\begin{frame}[t]{Sinusoid data movie}
{\tt https://www.youtube.com/watch?v=i-f4VqPPeGg}
\end{frame}

@@

\begin{frame}[t]{Recommendations}
It is infeasable to apply this approach wholesale.
Instead I would use it on small subsets of imaging that are of specific interest.
\end{frame}
